---
title: "2. MethyLight marker and threshold selection"
author: "Chiara Herzog"
date: April 4, 2022
output:
  rmarkdown::html_document:
    theme: cosmo
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggsci)
library(pROC)
library(patchwork)
library(ggrepel)
library(gridExtra)
library(broom)

# Set colours and themes
font <- theme_set(
  theme(text = element_text(family = "Guardian Sans")
  )
)
cols <- pal_jama(alpha = 0.8)(7)

# Read in data
load("../0-data/WID-qEC_manuscript_data.Rdata")
load("../0-data/tcga.Rdata")
load("../0-data/tcga_pvals.Rdata")

# Read in function
source("../1-source-codes/plot_roc.R")
```

# Ranking performance of MethyLight markers by AUC

* AUC is calculated for each individual region

```{r fig.align = "center", fig.width = 4.5, fig.height = 6}
mldat <- mldat %>%
  filter(set == "FORECEE Pilot") %>%
  droplevels() %>%
  mutate(type = factor(type, levels = c("Control", "Endometrial cancer")))

dat <- mldat %>%
  select(10:46) %>%
  droplevels()

rocs <- data.frame(matrix(nrow = ncol(dat), ncol = 3))
for (i in 1:ncol(dat)){
  x <- roc(as.factor(mldat$type), dat[,i])
  rocs[i,2] <- x$auc[[1]]
  rocs[i,3] <- coords(x, "best", transpose = TRUE)[[1]]
}

rocs[,1] <- colnames(dat)
colnames(rocs) <- c("name", "auc", "threshold")
rocs <- rocs[order(rocs$auc, decreasing = TRUE),]

rocs[,1] <- gsub("\n", " ", rocs[,1])
colours <- ifelse(grepl("ZSCAN12", rocs[,1]) == TRUE | grepl("GYPC", rocs[,1]) == TRUE, "selected", "not selected")

rocs %>%
  ggplot(aes(auc,
             reorder(name, auc),
             colour = colours,
             size = colours)) +
  geom_point() +
  xlab("AUC") +
  ylab("")  +
  scale_colour_manual(values = c("gray60", cols[2]),
                      name = "") +
  scale_size_manual(values = c(1, 2.5),
                    name = "") +
  xlim(c(0.7,1)) +
  theme(legend.position = "top",
        panel.grid.major = element_blank(),
        axis.text.y = element_text(size = 8,
                                   hjust = 1),
        legend.key = element_rect(fill = alpha("white", 0.0)),
        panel.grid = element_blank())
```


## Beta values of index CpGs in top three regions in TCGA-EC data

* TCGA data was accessed from TCGA using the TCGAbiolinks package using the following code:

```{r, eval = FALSE, echo = TRUE}
library(TCGAbiolinks)
query <- GDCquery(project = "TCGA-UCEC",
                  data.category = "DNA Methylation",
                  platform = "Illumina Human Methylation 450",
                  sample.type = c("Primary Tumor",
                                  "Solid Tissue Normal"))

query_pheno <- GDCquery_clinic(project = "TCGA-UCEC",
                               type = "clinical")

dat <- query$results[[1]]
ind <- match(dat$cases.submitter_id, query_pheno$submitter_id)

tcga_pheno <- query_pheno[ind,]
save(tcga_pheno, file = "tcga_pheno.Rdata")
GDCdownload(query, method = "client", files.per.chunk = 10) # RIP internet
data <- GDCprepare(query)
save(data, file = "data.Rdata")

library(SummarizedExperiment)
beta <- assays(data)[[1]]
beta <- na.omit(beta)
save(beta, file = "beta.Rdata")
```

* Plots:

```{r fig.align = "center", fig.width = 7, fig.height = 3.5}
# df <- tcga %>%
#   nest(-cg) %>%
#   mutate(test = map(data, ~ tidy(t.test(beta ~ type, .x)))) %>%
#   unnest(test) %>%
#   mutate(padj = p.adjust(p.value, n = 378254)) %>% # number of CpGs in beta
#   mutate(p = paste0("p=",signif(padj,3))) %>%
#   select(cg, p)   %>%
#   mutate(cg = factor(cg))

tcga %>%
  ggplot(aes(x = type,
             y = beta,
             colour = type)) +
  geom_violin() +
  geom_jitter(size = 0.2)+
  geom_boxplot(width = 0.1) +
  scale_colour_manual(values = cols[c(1,2)],
                      name = "") +
  facet_wrap(~cg) +
  annotate("segment",
           x = 1, xend = 2,
           y = 1, yend = 1,
           size = 0.25,
           colour = "gray40") +
  xlab("") +
  ylab("beta in TCGA data") +
  theme_gray() +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        text = element_text(family = "Guardian Sans")) +
  geom_text(data = tcga_pvals,
            mapping = aes(x = 1.5, y = 1.03, label = p),
            colour = "gray40",
            size = 3.5)
```

## Threshold selection for top three MethyLight regions

* Showing the individual log PMR values for the three top regions above, as well as the sum of their values.
* The thresholds are calculated using Youden's index on the receiver operating characteristic.
* The two thresholds (0.03 and 0.63) are then consistently used throughout the manuscript.

```{r fig.align = "center", fig.width = 9, fig.height = 8}

mldat <- mldat %>%
  pivot_longer(c(`ZSCAN12\n(cg25060829 [ii])`, `GYPC\n(cg15768103 [ii])`, `GYPC\n(cg15975865 [i])`), names_to = "locus", values_to = "pmr") 

p1 <- mldat %>%
  ggplot(aes(x = type,
             y = log(pmr+0.00001),
             colour = type)) +
  geom_violin(width = 0.2) +
  geom_jitter(size = 0.5,
              aes(colour = type)) +
  facet_wrap(~locus) +
  xlab("") +
  ylab(expression(paste("log (PMR)", sep = ""))) + 
  scale_colour_manual(values = cols[c(1,2)],
                      name = "") +
  theme(axis.text.x = element_text(angle = 60,
                                   hjust = 1),
        panel.grid = element_blank(),
        legend.position = "none"
  ) 

p2 <- mldat %>%
  ggplot(aes(x = type,
             y = log(sum+0.00001),
             colour = type)) +
  geom_violin(width = 0.2) +
  geom_jitter(size = 0.5,
              aes(colour = type)) +
  xlab("") +
  ylab(expression(paste(atop(bold("WID™-qEC"), "log ("~Sigma~" PMR)")))) + 
  scale_colour_manual(values = cols[c(1,2)],
                      name = "") +
  theme(axis.text.x = element_text(angle = 60,
                                   hjust = 1),
        panel.grid = element_blank(),
        legend.position = "none")

p3 <- plot_roc(as.factor(mldat$type), mldat$sum, "sum(PMR)") +
  theme(panel.grid = element_blank())

design <- "
AAAAA
BBCCD
"

y <- roc(mldat$type, mldat$sum)
cut <- cutoffs <- coords(y, "best", best.method = "youden")

cut_round <- round(cut, 2)
colnames(cut_round) <- c("Threshold", "Specificity (%)", "Sensitivity (%)")
cut_round$`Specificity (%)` <- cut_round$`Specificity (%)`*100
cut_round$`Sensitivity (%)` <- cut_round$`Sensitivity (%)`*100
cut_round <- t(cut_round)
colnames(cut_round) <- c("# 1", "# 2")

p3 <- p3 +
  geom_point(aes(x = 1-cut$specificity[2],
                 y = cut$sensitivity[2]),
             shape = 21,
             size = 3,
             fill = cols[3]) +
  annotate("text",
           label = "Threshold 2",
           x = 1-cut$specificity[2]+0.17,
           y = cut$sensitivity[2]-0.07,
           colour = cols[3],
           size = 2.95) +
  geom_point(aes(x = 1-cut$specificity[1],
                 y = cut$sensitivity[1]),
             shape = 24,
             size = 3,
             fill = cols[6]) +
  annotate("text",
           label = "Threshold 1",
           x = 1-cut$specificity[1]+0.17,
           y = cut$sensitivity[1]+0.05,
           colour = cols[6], size = 2.95)

p1 + p2 + p3 +
  gridExtra::tableGrob(cut_round,
                       theme = ttheme_default(base_size = 8)) +
  plot_layout(design = design)
```


